
\chapter{Introduction}
\label{chap:introduction}

The hardware setup is a model of a miniaturised trailer truck powered with motors and sensors with the ability to move autonomously. This trailer truck test bed was designed by Mr. Pravin Dhake in \cite{dhake2007real}  and Dr. Robert Woodley in \cite{woodley1999testbed}.
The main thrust of this work is to give the trailer truck some visual information about its environment. This information is gathered from a single camera . The camera provides with images, that  are used to perceive the unknown dimension. A set of function libraries were developed which allows the user to develop control algorithms based on the visual information.

This test bed consists of a miniaturized trailer bed, a single board computer and software libraries such as the data acquisition library, the control library and a computer vision library.  The system also used, open source image processing libraries like OpenCV and Point cloud to facilitate the input, output and certain low level  routines in the work . These libraries have a dearth of documentation and support available that really facilitate real-time acquisition of images and their processing. The libraries are scalable and can be used in the control of the trailer truck.
 
There are many potential applications of the designed system including numerous military and automotive applications. Similar  robustness and reliability requirements are desired in this system.

The thesis is divided into two sections. First section describes all the previous hardware, upgrades and the corresponding specifications, The second dection setals the software libraries and their dependencies along with the  different functions available in the library and their usage.